Function to find association b/w n significant eigen genes of meta modules with Diagnosis, Diagnosis.Gender, ApoE

### 2017-10-09 code/Rmd/networkAssociationTest_metaModules.Rmd
Added markdown

---
title: "Function to find association b/w n significant eigen genes with Diagnosis, Diagnosis.Gender, ApoE"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

@@

knit2synapse::knitToFolderEntity(file = "./networkAssociationTest.Rmd",
                                 parentId = "syn10306449",
                                 entityName = 'Network Association Tests')
```

@@

thisFileName <- 'networkAssociationTest.Rmd'

@@

### Data download
```{r download.data, include=FALSE}
# Download data
MODULES = synTableQuery('SELECT * FROM syn10338156')@values
all.used.ids = c('syn10338156')

@@

MODULE.OVRLP = synTableQuery('SELECT * FROM syn10339153')@values
all.used.ids = c(all.used.ids, 'syn10339153')
MODULE.ENRICH = synTableQuery('SELECT * FROM syn10492048')@values
all.used.ids = c(all.used.ids, 'syn10492048')

MAGMA.ENRICH = synTableQuery('SELECT * FROM syn10380432')@values
all.used.ids = c(all.used.ids, 'syn10380432')
```

@@

# Function to get module overlap plots
getOverlapPlots <- function(imp.mods, MOD.MEMBERS, MODULE.OVRLP){
    mdl.ovrlp = matrix(0, length(unique(imp.mods$ModuleNameFull)),length(unique(imp.mods$ModuleNameFull)))
    rownames(mdl.ovrlp) = unique(imp.mods$ModuleNameFull)
    colnames(mdl.ovrlp) = unique(imp.mods$ModuleNameFull)
    for( i in rownames(mdl.ovrlp)){
      for( j in colnames(mdl.ovrlp)){
        mdl.ovrlp[i,j] = length(intersect(MOD.MEMBERS[[i]],MOD.MEMBERS[[j]]))/length(unique(c(MOD.MEMBERS[[i]],MOD.MEMBERS[[j]])))
      }
    }

@@

Dx = lapply(assoc.stat.diagnosis, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  })

p = ggplot(Dx, aes(x = ngenes, y = total.eff.sz, color = method)) + geom_point() + xlim(0,150)
p = p + facet_grid(.~brainRegion, scales = 'free_x') + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = as.factor(npc.sig), y = total.eff.sz)) + geom_boxplot()
p = p + facet_grid(.~brainRegion, scales = 'free_x') + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = method, y = total.eff.sz, color = total.fdr)) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('Total Effect Size')
p = p + scale_color_continuous(low = 'red', high = 'white')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = method, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')

writeLines('Number of significantly associated modules')
Dx %>%
  dplyr::group_by(brainRegion, method) %>%
  dplyr::summarise(nsig = sum(total.fdr <= 0.05)) %>%
  tidyr::spread(method, nsig) %>%
  kable()

writeLines('Percent of significantly associated modules')
Dx %>%
  dplyr::group_by(brainRegion, method) %>%
  dplyr::summarise(nsig = round(sum(total.fdr <= 0.05)/sum(total.fdr <= 1) * 100, 2)) %>%
  tidyr::spread(method, nsig) %>%
  kable()

writeLines('Percent of modules whose PC1 is significantly associated')
Dx %>%
  dplyr::group_by(brainRegion, method) %>%
  dplyr::summarise(nsig = round(sum(pc1.fdr <= 0.05)/sum(pc1.fdr <= 1) * 100, 2)) %>%
  tidyr::spread(method, nsig) %>%
  kable()
```
Top 5 modules per method
```{r Dx.1, fig.height=12, fig.width=12}
# Top 5 in each method
h = lapply(unique(Dx$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion, method) %>%
    dplyr::filter(total.fdr <= 0.05) %>%
    dplyr::top_n(5, total.eff.sz) %>%
    dplyr::top_n(5, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, Dx, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```
Top 5 modules any method
```{r Dx.2, fig.height=12, fig.width=12}
# Top 5 any method
h = lapply(unique(Dx$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion) %>%
    dplyr::filter(total.fdr <= 0.05) %>%
    dplyr::top_n(25, total.eff.sz) %>%
    dplyr::top_n(25, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)
  print(summary(factor(imp.mods$method)))

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, Dx, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```

@@

Dx.gender = lapply(assoc.stat.diagnosis.gender, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  })

p = ggplot(Dx.gender, aes(x = ngenes, y = total.eff.sz, color = method)) + geom_point() + xlim(0,150)
p = p + facet_grid(.~brainRegion, scales = 'free_x') + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = as.factor(npc.sig), y = total.eff.sz)) + geom_boxplot()
p = p + facet_grid(.~brainRegion, scales = 'free_x') + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = method, y = total.eff.sz, color = total.fdr)) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('Total Effect Size')
p = p + scale_color_continuous(low = 'red', high = 'white')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = method, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')

writeLines('Number of significantly associated modules')
Dx.gender %>%
  dplyr::group_by(brainRegion, method) %>%
  dplyr::summarise(nsig = sum(total.fdr <= 0.05)) %>%
  tidyr::spread(method, nsig) %>%
  kable()

writeLines('Percent of significantly associated modules')
Dx.gender %>%
  dplyr::group_by(brainRegion, method) %>%
  dplyr::summarise(nsig = round(sum(total.fdr <= 0.05)/sum(total.fdr <= 1) * 100, 2)) %>%
  tidyr::spread(method, nsig) %>%
  kable()

writeLines('Percent of modules whose PC1 is significantly associated')
Dx.gender %>%
  dplyr::group_by(brainRegion, method) %>%
  dplyr::summarise(nsig = round(sum(pc1.fdr <= 0.05)/sum(pc1.fdr <= 1) * 100, 2)) %>%
  tidyr::spread(method, nsig) %>%
  kable()
```
Top 5 modules per method
```{r Dx.gender.1, fig.height=12, fig.width=12}
# Top 5 in each method
h = lapply(unique(Dx.gender$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion, method) %>%
    dplyr::filter(total.fdr <= 0.05) %>%
    dplyr::top_n(5, total.eff.sz) %>%
    dplyr::top_n(5, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)

  if(dim(imp.mods)[1] != 0 ){
    getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
  }
}, Dx.gender, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```
Top 5 modules any method
```{r Dx.gender.2, fig.height=12, fig.width=12}
# Top 5 any method
h = lapply(unique(Dx.gender$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion) %>%
    dplyr::filter(total.fdr <= 0.05) %>%
    dplyr::top_n(25, total.eff.sz) %>%
    dplyr::top_n(25, pc1.eff.sz) %>%
    dplyr::filter(brainRegion == br)

  if(dim(imp.mods)[1] != 0 ){
    print(summary(factor(imp.mods$method)))
    getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
  }
}, Dx.gender, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
# draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```

### Associaiton with ApoE
```{r ApoE, fig.height=5, fig.width=10}
# Find assocition b/w ApoE category with n eigen genes
FORMULA = as.formula('PC~BrainRegion.ApoE+Batch+RIN+PCT_INTRONIC_BASES+PMI+PCT_INTERGENIC_BASES+AOD+PCT_PF_READS_ALIGNED+Gender')
assoc.stat.apoe = MODULES %>%
  dplyr::filter(GeneID %in% rownames(EXPR)) %>%
  plyr::dlply(.(ModuleName, brainRegion), .fun = findAssociationStatistics, 
              COVARIATES, EXPR, FORMULA,
              .parallel = T, .paropts = list(.packages = c('dplyr', 'data.table')))

ApoE = lapply(assoc.stat.apoe, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  })

p = ggplot(ApoE, aes(x = ngenes, y = total.eff.sz, color = method)) + geom_point() + xlim(0,150)
p = p + facet_grid(.~brainRegion, scales = 'free_x') + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(ApoE, aes(x = as.factor(npc.sig), y = total.eff.sz)) + geom_boxplot()
p = p + facet_grid(.~brainRegion, scales = 'free_x') + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(ApoE, aes(x = method, y = total.eff.sz, color = total.fdr)) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('Total Effect Size')
p = p + scale_color_continuous(low = 'red', high = 'white')
p + theme(legend.position = 'top')

p = ggplot(ApoE, aes(x = method, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + facet_grid(.~brainRegion, scales = 'free_x')  + xlab('Method') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')
```

### Select information rich module (based on Dx association, MAGMA enrichment, dbgap enrichment and AD related pathway enrichment)
```{r imp.mdl, fig.height=7, fig.width=12}
ad.magma = MAGMA.ENRICH %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(fdr = p.adjust(P),
                    rnk = rank(rank(-log10(fdr))+rank(BETA)),
                    rnk = rnk/max(rnk)) %>%
      dplyr::arrange(desc(rnk))%>%
      dplyr::group_by(SET) %>%
      dplyr::top_n(1, rnk) %>%
      dplyr::top_n(1, -log10(fdr))
  })

ad.dbgap = MODULE.ENRICH %>%
  dplyr::filter(geneSet == 'dbgap', category == 'Alzheimer.Disease') %>%
  unique() %>%
  dplyr::left_join(Dx %>% dplyr::select(ModuleName, brainRegion, ModuleNameFull, method)) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(fdr = p.adjust(fisherPval),
                    rnk = rank(rank(-log10(fdr))+rank(fisherOR)),
                    rnk = rnk/max(rnk)) %>%
      dplyr::arrange(desc(rnk)) %>%
      dplyr::group_by(ModuleNameFull) %>%
      dplyr::top_n(1, rnk) %>%
      dplyr::top_n(1, -log10(fdr))
  })

other.ad.pathways = c("alzheimer_disease", "Alzheimer.s.disease_Homo.sapiens_hsa05010", 
                      "Alzheimers.Disease_Homo.sapiens_WP2059",
                      "Alzheimer.disease.amyloid.secretase.pathway_Homo.sapiens_P00003",
                      "Alzheimer.s_disease", "Alzheimer.Disease")
ad.others = MODULE.ENRICH %>%
  dplyr::filter(geneSet != 'dbgap', category %in% other.ad.pathways) %>%
  unique() %>%
  dplyr::left_join(Dx %>% dplyr::select(ModuleName, brainRegion, ModuleNameFull, method)) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(fdr = p.adjust(fisherPval),
                    rnk = rank(rank(-log10(fdr))+rank(fisherOR)),
                    rnk = rnk/max(rnk)) %>%
      dplyr::arrange(desc(rnk))%>%
      dplyr::group_by(ModuleNameFull) %>%
      dplyr::top_n(1, rnk) %>%
      dplyr::top_n(1, -log10(fdr))
  })

# Combine and rank all the information
scr1 = Dx %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(rnk = rank(rank(total.eff.sz) + rank(-log10(total.fdr))),
                    rnk = rnk/max(rnk)) %>%
      dplyr::arrange(desc(rnk))%>%
      dplyr::group_by(ModuleNameFull) %>%
      dplyr::top_n(1, rnk) %>%
      dplyr::top_n(1, -log10(pc1.fdr))
  }) %>%
  dplyr::mutate(Dx = rank(rnk),
                Dx = Dx/max(Dx)) %>%
  dplyr::select(ModuleNameFull, brainRegion, method, ModuleName, Dx)

scr2 = ad.magma %>%
  dplyr::mutate(magma = rank(rnk),
                magma = magma/max(magma)) %>%
  dplyr::rename(ModuleNameFull = SET) %>%
  dplyr::select(ModuleNameFull, brainRegion, method, magma)

scr3 = ad.dbgap %>%
  dplyr::mutate(dbgap = rank(rnk),
                dbgap = dbgap/max(dbgap)) %>%
  dplyr::select(ModuleNameFull, ModuleName, brainRegion, method, dbgap)

scr4 = ad.others %>%
  dplyr::mutate(other = rank(rnk),
                other = other/max(other)) %>%
  dplyr::select(ModuleNameFull, ModuleName, brainRegion, method, other)

all.rank = plyr::join_all(list(scr1, scr2, scr3, scr4), type = 'full')
all.rank[is.na(all.rank)] = 0
all.rank = all.rank %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x = x %>%
      dplyr::mutate(rnk = rank(Dx+magma+dbgap+other),
                    rnk = rnk/max(rnk))
    print(paste('BrainRegion:', unique(x$brainRegion)))
    print(paste('Correlation with Dx', round(cor(x$rnk, x$Dx),2)))
    print(paste('Correlation with magma', round(cor(x$rnk, x$magma),2)))
    print(paste('Correlation with dbgap', round(cor(x$rnk, x$dbgap),2)))
    print(paste('Correlation with others', round(cor(x$rnk, x$other),2)))
    return(x)
  })

ggplot(all.rank, aes(x = Dx, y = magma))+geom_point()+facet_grid(.~brainRegion)

ggplot(all.rank, aes(x = dbgap, y = other))+geom_point()+facet_grid(.~brainRegion)+geom_smooth(method = 'lm')
ggplot(all.rank, aes(x = dbgap, y = magma))+geom_point()+facet_grid(.~brainRegion)+geom_smooth(method = 'lm')

ggplot(all.rank, aes(x = other, y = magma))+geom_point()+facet_grid(.~brainRegion)+geom_smooth(method = 'lm')

ggplot(tmp, aes(x = method, y = rnk)) + geom_boxplot() + facet_grid(.~brainRegion) + theme(axis.text.x = element_text(angle=45, hjust = 1)) + coord_flip() + geom_jitter()

```
Top 5 modules per method
```{r imp.mdl.1, fig.height=12, fig.width=12}
# Top 5 in each method
h = lapply(unique(all.rank$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion, method) %>%
    dplyr::top_n(5, rnk) %>%
    dplyr::top_n(5, Dx) %>%
    dplyr::filter(brainRegion == br)

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, all.rank, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```
Top 5 modules any method
```{r imp.mdl.2, fig.height=12, fig.width=12}
# Top 5 any method
h = lapply(unique(all.rank$brainRegion), function(br, Dx, MOD.MEMBERS, MODULE.OVRLP){
  imp.mods = Dx %>%
    dplyr::group_by(brainRegion) %>%
    dplyr::top_n(25, rnk) %>%
    dplyr::top_n(25, Dx) %>%
    dplyr::filter(brainRegion == br)
  print(summary(factor(imp.mods$method)))

  getOverlapPlots(imp.mods, MOD.MEMBERS, MODULE.OVRLP)
}, all.rank, MOD.MEMBERS, MODULE.OVRLP)
draw(h[[1]], heatmap_legend_side='left')
draw(h[[2]], heatmap_legend_side='left')
draw(h[[3]], heatmap_legend_side='left')
draw(h[[4]], heatmap_legend_side='left')
draw(h[[5]], heatmap_legend_side='left')
draw(h[[6]], heatmap_legend_side='left')
draw(h[[7]], heatmap_legend_side='left')
```

### Modules missing enrichment scores
```{r missing, include=FALSE}
writeLines('Modules missing enrich enrichment scores')
filter(Dx, ModuleNameFull %in% unique(setdiff(Dx$ModuleNameFull, MODULE.ENRICH$ModuleNameFull))) %>%
  dplyr::select(ModuleNameFull) %>% unique() %>% unlist() %>% paste(collapse = ', ') %>% print()

writeLines('Modules missing MAGMA enrichment scores')
filter(Dx, ModuleNameFull %in% unique(setdiff(Dx$ModuleNameFull, MAGMA.ENRICH$Set))) %>%
  dplyr::select(ModuleNameFull) %>% unique() %>% unlist() %>% paste(collapse = ', ') %>% print()
```

### Store results in synapse
```{r syn.store, include=FALSE}
write.table(Dx, file = 'AssociaitonStatWithDx.tsv', sep = '\t', quote = F, row.names = F)
OBJ = File('AssociaitonStatWithDx.tsv', name = 'Association with Dx', parentId = 'syn10831307')
OBJ = synStore(OBJ, used = all.used.ids, executed = thisFile, activityName = activityName)

@@

write.table(ApoE, file = 'AssociaitonStatWithApoE4.tsv', sep = '\t', quote = F, row.names = F)
OBJ2 = File('AssociaitonStatWithApoE4.tsv', name = 'Association with ApoE4', parentId = 'syn10831307')
OBJ2 = synStore(OBJ2, used = all.used.ids, executed = thisFile, activityName = activityName)

save(list = c('assoc.stat.diagnosis', 'assoc.stat.diagnosis.gender', 'assoc.stat.apoe'), file = 'AllAssociationResults.RData')
OBJ3 = File('AllAssociationResults.RData', name = 'All Assocition Statistics (RData)', parentId = 'syn10831307')
OBJ3 = synStore(OBJ3, used = all.used.ids, executed = thisFile, activityName = activityName)
```

>>>>

---
title: "Function to find association b/w n significant eigen genes of meta modules with Diagnosis, Diagnosis.Gender, ApoE"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

@@

knit2synapse::knitToFolderEntity(file = "./networkAssociationTest_metaModules.Rmd",
                                 parentId = "syn10306449",
                                 entityName = 'Network Association Tests (aggregated modules)')
```

@@

thisFileName <- 'networkAssociationTest_metaModules.Rmd'

@@

### Data download
```{r download.data, include=FALSE}
# Download data
MODULES = synTableQuery('SELECT * FROM syn10915669')@values
all.used.ids = c('syn10915669')

@@

# MODULE.OVRLP = synTableQuery('SELECT * FROM syn10339153')@values
# all.used.ids = c(all.used.ids, 'syn10339153')
# MODULE.ENRICH = synTableQuery('SELECT * FROM syn10492048')@values
# all.used.ids = c(all.used.ids, 'syn10492048')
# 
# MAGMA.ENRICH = synTableQuery('SELECT * FROM syn10380432')@values
# all.used.ids = c(all.used.ids, 'syn10380432')
```

@@

# Function to get module overlap plots
getOverlapPlots <- function(imp.mods, MOD.MEMBERS){
    mdl.ovrlp = matrix(0, length(unique(imp.mods$ModuleNameFull)),length(unique(imp.mods$ModuleNameFull)))
    rownames(mdl.ovrlp) = unique(imp.mods$ModuleNameFull)
    colnames(mdl.ovrlp) = unique(imp.mods$ModuleNameFull)
    for( i in rownames(mdl.ovrlp)){
      for( j in colnames(mdl.ovrlp)){
        mdl.ovrlp[i,j] = length(intersect(MOD.MEMBERS[[i]],MOD.MEMBERS[[j]]))/length(unique(c(MOD.MEMBERS[[i]],MOD.MEMBERS[[j]])))
        fisherEnrichment()
      }
    }

@@

Dx = lapply(assoc.stat.diagnosis, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  }) %>%
  dplyr::mutate(total.sig = total.fdr <= 0.05)

p = ggplot(Dx, aes(x = ngenes, y = total.eff.sz, color = brainRegion)) + geom_point()
p = p + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = as.factor(npc.sig), y = total.eff.sz, color = brainRegion)) + geom_boxplot()
p = p + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = brainRegion, y = total.eff.sz)) + geom_boxplot() + geom_jitter(aes(color = total.sig))
p = p + coord_flip() + xlab('Brain Region') + ylab('Total Effect Size')
# p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')

p = ggplot(Dx, aes(x = brainRegion, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + xlab('Brain Region') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')

writeLines('Number of significantly associated modules')
Dx %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = sum(total.fdr <= 0.05),
                   ntotal = sum(total.fdr <= 1)) %>%
  kable()

writeLines('Percent of significantly associated modules')
Dx %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = round(sum(total.fdr <= 0.05)/sum(total.fdr <= 1) * 100, 2)) %>%
  kable()

writeLines('Percent of modules whose PC1 is significantly associated')
Dx %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = round(sum(pc1.fdr <= 0.05)/sum(pc1.fdr <= 1) * 100, 2)) %>%
  kable()
```

@@

Dx.gender = lapply(assoc.stat.diagnosis.gender, function(x) x$assoc.stat) %>% 
  data.table::rbindlist(use.names = T, fill = T, idcol = 'ModuleName.BrainRegion') %>%
  tidyr::separate(ModuleName.BrainRegion, c('ModuleName', 'brainRegion'), sep = '\\.') %>%
  dplyr::left_join(MODULES %>%
                     dplyr::select(Module, method, ModuleName, brainRegion, ModuleNameFull) %>%
                     unique()) %>%
  plyr::ddply(.(brainRegion), .fun = function(x){
    x %>%
      dplyr::mutate(total.fdr = p.adjust(total.pval))
  }) %>%
  dplyr::mutate(total.sig = total.fdr <= 0.05)

p = ggplot(Dx.gender, aes(x = ngenes, y = total.eff.sz, color = brainRegion)) + geom_point()
p = p + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = as.factor(npc.sig), y = total.eff.sz, color = brainRegion)) + geom_boxplot()
p = p + xlab('No. of genes') + ylab('Total Effect Size')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = brainRegion, y = total.eff.sz)) + geom_boxplot() + geom_jitter(aes(color = total.sig))
p = p + coord_flip() + xlab('Brain Region') + ylab('Total Effect Size')
# p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')

p = ggplot(Dx.gender, aes(x = brainRegion, y = pc1.eff.sz, color = -log10(pc1.fdr))) + geom_boxplot() + geom_jitter()
p = p + coord_flip() + xlab('Brain Region') + ylab('PC1 Effect Size')
p = p + scale_color_continuous(low = 'white', high = 'red')
p + theme(legend.position = 'top')

writeLines('Number of significantly associated modules')
Dx.gender %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = sum(total.fdr <= 0.05),
                   ntotal = sum(total.fdr <= 1)) %>%
  kable()

writeLines('Percent of significantly associated modules')
Dx.gender %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = round(sum(total.fdr <= 0.05)/sum(total.fdr <= 1) * 100, 2)) %>%
  kable()

writeLines('Percent of modules whose PC1 is significantly associated')
Dx.gender %>%
  dplyr::group_by(brainRegion) %>%
  dplyr::summarise(nsig = round(sum(pc1.fdr <= 0.05)/sum(pc1.fdr <= 1) * 100, 2)) %>%
  kable()
```

### Enrichment for differential expression
```{r Dx.enrich}
# Get differential expression gene list from synapse
load(synGet('syn10496554')@filePath)
all.used.ids = c(all.used.ids, 'syn10496554')
names(amp.ad.de.geneSets) = gsub('CER','CBE', names(amp.ad.de.geneSets))

# Perform enrichment analysis
MOD.MEMBERS = plyr::dlply(MODULES, .(ModuleNameFull), .fun = function(x) unique(x$external_gene_name))
Dx.enrich = plyr::ldply(unique(MODULES$brainRegion), .fun = function(x){
  modMembers = MOD.MEMBERS[grep(x, names(MOD.MEMBERS))]
  deGenes = amp.ad.de.geneSets[grep(x, names(amp.ad.de.geneSets))]
  backgroundGenes = unique(MODULES$external_gene_name)

  enrichResults = plyr::ldply(modMembers, .fun = function(y, deGenes, backgroundGenes){
    plyr::ldply(deGenes, .fun = function(z, y, backgroundGenes){
      fisherEnrichment(y,z,backgroundGenes)
    }, y, backgroundGenes, .id = 'Dx.GeneSet')
  }, deGenes, backgroundGenes, .id = 'ModuleNameFull') %>%
    dplyr::mutate(fdr = p.adjust(pval))
}, .id = 'brainRegion')
```
### Enrichment for gene sets
```{r pathway.enrich}
# Download all related genesets from synapse
load(synGet('syn4867851')@filePath)
all.used.ids = c(all.used.ids, 'syn4867851')

# Download cell type specific genesets
load(synGet('syn9818080')@filePath)
all.used.ids = c(all.used.ids, 'syn9818080')

backgroundGenes = unique(MODULES$external_gene_name)
geneSet.ref = c(GeneSets[c('KEGG_2016', 'GO_Biological_Process_2015', 'Reactome_2016', 'WikiPathways_2016')],
                list(CellType = CellTypeList) )%>%
  lapply(function(x){
    lapply(x, function(y){
      sapply(y, function(y) str_split(y, '\\,')[[1]][1])
    })
  }) %>%
  CovariateAnalysis::filterGeneSets(backgroundGenes) %>%
  do.call(c,.)

# Perform enrichment analysis
Pathway.enrich = plyr::ddply(MODULES, .(ModuleNameFull), .fun = function(mod, geneSetRef, backgroundGenes){
  modMembers = unique(mod$external_gene_name)

  enrichResults = plyr::ldply(geneSetRef, .fun = function(y, modMembers, backgroundGenes){
    fisherEnrichment(y, modMembers, backgroundGenes)
  }, modMembers, backgroundGenes, .id = 'GeneSetName', .parallel = T, 
  .paropts = list(.packages = c('CovariateAnalysis'))) %>%
    dplyr::mutate(fdr = p.adjust(pval))
}, 
.id = 'ModuleNameFull', geneSet.ref, backgroundGenes)
```

### Store results in synapse
```{r syn.store, include=FALSE, eval = F}
write.table(Dx, file = 'AssociaitonStatWithDx.tsv', sep = '\t', quote = F, row.names = F)
OBJ = File('AssociaitonStatWithDx.tsv', name = 'Association with Dx', parentId = 'syn10831307')
OBJ = synStore(OBJ, used = all.used.ids, executed = thisFile, activityName = activityName)

@@

write.table(Pathway.enrich, file = 'PathwayEnrichment.tsv', sep = '\t', quote = F, row.names = F)
OBJ2 = File('PathwayEnrichment.tsv', name = 'Pathway Enrichment', parentId = 'syn10831307')
OBJ2 = synStore(OBJ2, used = all.used.ids, executed = thisFile, activityName = activityName)

save(list = c('assoc.stat.diagnosis', 'assoc.stat.diagnosis.gender'), file = 'AllAssociationResults.RData')
OBJ3 = File('AllAssociationResults.RData', name = 'All Assocition Statistics (RData)', parentId = 'syn10831307')
OBJ3 = synStore(OBJ3, used = all.used.ids, executed = thisFile, activityName = activityName)

write.table(Dx.enrich, file = 'DiagnosisEnrichment.tsv', sep = '\t', quote = F, row.names = F)
OBJ4 = File('DiagnosisEnrichment.tsv', name = 'Diagnosis Enrichment', parentId = 'syn10831307')
OBJ4 = synStore(OBJ4, used = all.used.ids, executed = thisFile, activityName = activityName)
```
